\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{geometry}
\geometry{margin=2.5cm}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}

\title{\textbf{Paper A-01: Earth Unified Environmental Transport:\\ A Constraint-Driven Flux Dynamics (CDFD) Perspective}}
\author{Steve Bico Mujjabi, MD\\
\small Independent Researcher\\
\small ORCID: 0009-0001-0556-5516}
\date{May 2026}

\begin{document}
\maketitle

\begin{abstract}
This opening Earth-systems paper states the Part A modelling frame. It treats
environmental transport as a CDFD/AFL hypothesis in which heat, water,
nutrients, sediment, pollutants, and biological energy are represented as
domain-specific fluxes $\Phi$ moving against constraints $C$ through surfaces
with responsiveness $S$ and structural memory $M_s$. The release-local runtime
outputs are internal diagnostics and tests that could break the mapping, not field evidence
of climate, hydrological, ecological, or geological claims.
\end{abstract}

\paragraph{Greek-symbol convention.}
Unless a manuscript states a tighter equation-local meaning, Greek symbols are
local CDFD variables or coefficients, not universal constants. Here $\Phi$
denotes driving flux, $\Psi_s$ denotes the adaptive operating ratio, and
$\Lambda$ denotes the Life Number where used. The coefficients $\alpha$,
$\beta$, and $\gamma$ denote accumulation or reinforcement, relaxation or decay,
and diffusion, coupling, or redistribution. Subscripts restrict coefficients to
the named pathway, domain, or experiment.

\section{Earth-System Transport Frame}

Earth systems are coupled transport networks. Atmospheric circulation moves
heat and moisture; oceans redistribute heat and salinity; rivers move water and
sediment; soils regulate nutrients and organic matter; ecosystems route biomass
and information through trophic structure. CDFD/AFL does not replace the
standard equations in these fields. It supplies a common audit layer for asking
where flow rises, where capacity tightens, where responsiveness is lost, and
where memory keeps a system from returning to its earlier state.

\section{Candidate Variable Map}

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
CDFD term & Earth-system reading & Example proxy \\
\midrule
$\Phi$ & environmental drive or throughput & heat flux, runoff, nutrient load \\
$C$ & active constraint or capacity burden & aridity, channel resistance, toxicity \\
$S$ & responsiveness or buffering & biodiversity, soil structure, wetland storage \\
$M_s$ & structural memory & degradation, hysteresis, legacy contamination \\
$\Psi_s$ & operating ratio & balanced, constrained, or overloaded regime \\
\bottomrule
\end{tabular}
\end{table}

This map is hypothesis-level until the proxies, scales, datasets, and failure
conditions are specified for a given Earth-system question.

\section{Part A Sequence}

The Part A papers move from broad environmental transport into atmosphere,
ocean, hydrology, soil, ecology, biodiversity, pollution, agriculture,
sustainability, collapse, synthesis, plate tectonics, and volcanology. The
sequence is ordered to start with transport structure, then add specific
environmental constraints and failure modes.

\section{Measurement Layer}

The first job in any domain is to choose proxies before drawing conclusions. $\Phi$ may be a rate of material, energy, information, capital, or signal flow. $C$ is the bottleneck that slows or redirects that flow. $S$ measures the system's ability to reconfigure under stress, and $M_s$ records the past load that still changes present behavior. A useful application gives those four terms observable definitions, a time scale, and a failure condition. If a standard domain model explains the same data more cleanly, the CDFD/AFL mapping should give way.

\section{Current Runtime Diagnostic}

The diagnostic script \texttt{Part\_E\_Synthesis/supplementary/run\_partiv\_discovery.py}
includes selected Earth-system adapter rows for climate and atmosphere, ocean
transport, hydrological flow, biodiversity, and pollution. The accompanying
network-cascade diagnostic checks finite-value behavior in an overload and
memory-locking stress test. I use those rows as a runtime audit, not as a
replacement for Earth-system data.

\section{Falsification Targets}

The Earth-systems framing weakens if a domain cannot define measurable
proxies for $\Phi$, $C$, $S$, and $M_s$, if the proposed operating ratio fails
to predict overload or recovery behavior, or if a standard domain model explains
the same data without any added value from the CDFD/AFL mapping.

\section{Conclusion}

Part A uses CDFD/AFL as a shared transport language for Earth systems. The
framework is useful only when it sharpens measurable hypotheses and failure
conditions. It remains separate from operational climate, hydrological,
ecological, agricultural, or geological advice.

\section*{Conflict of Interest}
The author declares no conflict of interest.

\section*{Data Availability}
Release-local runtime diagnostics are available under each Part's \texttt{outputs/} folder, with release-wide synthesis artifacts under \texttt{Part\_E\_Synthesis/supplementary/}, \texttt{Part\_E\_Synthesis/outputs/}, and \texttt{Part\_E\_Synthesis/figures/}.

\section{Reference Layer}
\bibliographystyle{plain}
\bibliography{universal_afl_references}
\end{document}
