\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{geometry}
\geometry{margin=2.5cm}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{criterion}{Criterion}




\title{\textbf{Paper A-02: Earth Atmospheric Circulation:\\ A Constraint-Driven Flux Dynamics (CDFD) Perspective}}
\author{Steve Bico Mujjabi, MD\\
\small Independent Researcher\\
\small ORCID: 0009-0001-0556-5516}
\date{May 2026}


\newcommand{\EarthSuppPath}{Part\_A\_Earth\_Systems/supplementary/supplementary\_earth\_}
\newcommand{\EngSuppPath}{Part\_B\_Engineered\_Systems/supplementary/supplementary\_eng\_}
\newcommand{\SocSuppPath}{Part\_C\_Socioeconomic\_Systems/supplementary/supplementary\_soc\_}
\newcommand{\DomainSuppPath}{Part\_D\_Domain\_Applications/supplementary/supplementary\_dom\_}
\begin{document}
\maketitle

\begin{abstract}
This paper treats atmospheric circulation as a CDFD/AFL modelling hypothesis.
Solar heating, latent heat, water vapour, aerosols, and momentum are read as
driving fluxes $\Phi$ moving through pressure gradients, rotation, stratified
layers, topography, and boundary-layer drag. The purpose is not to replace
meteorology or climate dynamics. The purpose is to name the measurable flux,
constraint, responsiveness, and memory terms that make atmospheric overload or
stagnation testable.
\end{abstract}

\section{Series Context}

Part I sets the flow--constraint language used here \cite{MujjabiPartI2026}.
Part II develops the Life Number and the origins-of-life use of the same notation
\cite{MujjabiPartII2026}. Part III applies AFL to biology and medicine
\cite{MujjabiPartIII2026}. Part IV asks whether the same audit language clarifies
cross-domain measurement without turning analogy into evidence.

\paragraph{Notation.}
In this paper, $C$ names the active atmospheric constraint or capacity burden.
$\Phi$ is the driving flow, $S$ is the surface or circulation response, and
$M_s$ is retained structural memory. $\Lambda$ appears only as the Part II Life
Number reference \cite{MujjabiPartII2026}; it is not an atmospheric variable in
this manuscript.

\section{Atmospheric Transport Frame}

The atmosphere is a moving constraint network. Uneven heating creates pressure
gradients; rotation bends flow; mountains, land--sea contrast, cloud formation,
surface roughness, and radiative forcing reshape the pathways available to air
and moisture. CDFD/AFL treats those pathways as an audit surface: where does
flux rise, where does the atmosphere lose room to adjust, and where does memory
from prior heat, moisture, or aerosol loading still shape the next state?

\section{Candidate Variable Map}

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
CDFD term & Atmospheric reading & Example proxy \\
\midrule
$\Phi$ & heat, moisture, aerosol, or momentum flux & radiative flux, latent heat, wind stress \\
$C$ & active circulation constraint & inversion strength, topography, drag, stability \\
$S$ & responsiveness & convective adjustment, mixing, storm-track mobility \\
$M_s$ & structural memory & soil moisture, sea-surface heat, aerosol persistence \\
$\Psi_s$ & operating ratio & ventilated, stagnant, amplified, or overloaded regime \\
\bottomrule
\end{tabular}
\end{table}

\section{Overload and Stagnation}

An atmospheric overload claim is meaningful only when the flux and constraint
are named together. A heat-wave example might use radiative and sensible heat
flux as $\Phi$, boundary-layer stability and weak advection as $C$, convective
mixing as $S$, and stored soil or ocean heat as $M_s$. A pollution-stagnation
example might use emissions as $\Phi$, inversion strength as $C$, turbulent
mixing as $S$, and accumulated aerosol load as $M_s$. These mappings remain
hypotheses until compared with atmospheric observations and standard models.

\section{Measurement Layer}

The measurable layer comes first. A useful atmospheric application gives
$\Phi$, $C$, $S$, and $M_s$ observable proxies, a spatial grid or network, a time
window, and a failure condition such as persistent stagnation, blocked moisture
transport, convective inhibition, or extreme heat amplification. If conventional
fluid-dynamical variables explain the event without added value from CDFD/AFL,
the mapping adds no release-level claim.

\section{Current Runtime Diagnostic}

The Part E diagnostic run includes a climate/atmosphere adapter row and the
universal cascade stress test. These files check finite runtime behavior and
regime reporting in the current CDFD Runtime \cite{CDFDRuntime2026}. They are a
model audit for later measurement, not a replacement for atmospheric data.

\section{Tests That Can Break the Mapping}

The atmospheric mapping weakens if measured flux and constraint proxies do not
track stagnation, amplification, or recovery; if the same event is explained
more cleanly by established atmospheric diagnostics; or if the CDFD/AFL terms
cannot be separated from ordinary variable names in a reproducible analysis.

\section{Conclusion}

Atmospheric circulation can be read as constrained transport when the variables
are tied to observations. The CDFD/AFL contribution is a common audit language
for flux, bottleneck, response, and memory. Its value depends on whether that
language sharpens measurable atmospheric questions.

\section*{Conflict of Interest}
The author declares no conflict of interest.

\section*{Data Availability}
Release-local runtime diagnostics are available under each Part's
\texttt{outputs/} folder, with release-wide synthesis artifacts under
\texttt{Part\_E\_Synthesis/supplementary/},
\texttt{Part\_E\_Synthesis/outputs/}, and
\texttt{Part\_E\_Synthesis/figures/}.

\section{Reference Layer}

\bibliographystyle{plain}
\bibliography{universal_afl_references}
\end{document}
