\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{geometry}
\geometry{margin=2.5cm}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{criterion}{Criterion}




\title{\textbf{Paper A-03: Earth Ocean Currents:\\ A Constraint-Driven Flux Dynamics (CDFD) Perspective}}
\author{Steve Bico Mujjabi, MD\\
\small Independent Researcher\\
\small ORCID: 0009-0001-0556-5516}
\date{May 2026}


\newcommand{\EarthSuppPath}{Part\_A\_Earth\_Systems/supplementary/supplementary\_earth\_}
\newcommand{\EngSuppPath}{Part\_B\_Engineered\_Systems/supplementary/supplementary\_eng\_}
\newcommand{\SocSuppPath}{Part\_C\_Socioeconomic\_Systems/supplementary/supplementary\_soc\_}
\newcommand{\DomainSuppPath}{Part\_D\_Domain\_Applications/supplementary/supplementary\_dom\_}
\begin{document}
\maketitle

\begin{abstract}
This paper treats ocean currents as a CDFD/AFL modelling hypothesis. Wind
stress, buoyancy forcing, heat, salt, and mass transport are read as fluxes
$\Phi$ moving through basin geometry, stratification, bathymetry, mixing limits,
and freshwater constraints $C$. The claim is not that CDFD/AFL replaces
physical oceanography. The claim is that the notation can organise measurable
questions about transport, bottlenecks, responsiveness, and long memory in the
ocean system.
\end{abstract}

\section{Series Context}

Part I sets the flow--constraint language used here \cite{MujjabiPartI2026}.
Part II develops the Life Number and the origins-of-life use of the same notation
\cite{MujjabiPartII2026}. Part III applies AFL to biology and medicine
\cite{MujjabiPartIII2026}. Part IV uses the language as a cross-domain stress
test, with measurement kept separate from analogy.

\paragraph{Notation.}
In this paper, $C$ names the active ocean constraint or capacity burden.
$\Phi$ is the driving transport, $S$ is the ocean response, and $M_s$ is retained
structural memory. $\Lambda$ is mentioned only where the Part II Life Number is
being referenced \cite{MujjabiPartII2026}.

\section{Ocean Transport Frame}

The ocean stores and redistributes heat, salt, carbon, nutrients, and momentum.
Its flow is constrained by basin walls, seafloor geometry, density structure,
wind forcing, sea ice, freshwater input, and mixing depth. In the CDFD/AFL
reading, a current system is not only a velocity field; it is a flow path with
capacity, inertia, adaptive response, and memory.

\section{Candidate Variable Map}

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
CDFD term & Ocean-current reading & Example proxy \\
\midrule
$\Phi$ & ocean transport or forcing & Sverdrup transport, heat flux, freshwater flux \\
$C$ & active constraint & stratification, bathymetry, sill depth, friction \\
$S$ & responsiveness & mixing, eddy compensation, current-path adjustment \\
$M_s$ & structural memory & heat storage, salinity anomaly, overturning inertia \\
$\Psi_s$ & operating ratio & balanced, constrained, stalled, or amplified transport \\
\bottomrule
\end{tabular}
\end{table}

\section{Memory and Circulation Change}

Ocean memory is central to the mapping. Heat stored below the mixed layer,
persistent salinity anomalies, ice-ocean feedback, and overturning inertia can
make present circulation depend on prior forcing. In CDFD/AFL terms, high
$M_s$ means the current state carries a record of earlier flux and constraint.
That does not by itself prove a new mechanism; it gives a way to state what
must be measured.

\section{Measurement Layer}

A useful ocean application specifies the current system, the forcing window,
the proxy for $\Phi$, the constraint term $C$, the response term $S$, and the
memory term $M_s$. Candidate failure conditions include persistent transport
weakening, blocked heat redistribution, salinity-driven stratification, or a
shift in overturning regime. The mapping has value only if it improves
measurement or interpretation beyond ordinary oceanographic diagnostics.

\section{Current Runtime Diagnostic}

The Part E diagnostic run includes an ocean-transport adapter row and the
universal cascade stress test. These outputs check finite runtime behavior in
the current CDFD Runtime \cite{CDFDRuntime2026}. They do not validate any ocean
claim; they record model behavior that later data work can test.

\section{Tests That Can Break the Mapping}

The ocean-current mapping weakens if the chosen proxies for flux, constraint,
responsiveness, and memory do not track observed transport changes; if a
standard ocean model explains the same behavior more cleanly; or if the CDFD
terms cannot be tied to reproducible ocean data.

\section{Conclusion}

Ocean currents are a natural test case for flow, constraint, response, and
memory. CDFD/AFL is useful here only as a disciplined measurement language. The
release-level claim remains limited to that role.

\section*{Conflict of Interest}
The author declares no conflict of interest.

\section*{Data Availability}
Release-local runtime diagnostics are available under each Part's
\texttt{outputs/} folder, with release-wide synthesis artifacts under
\texttt{Part\_E\_Synthesis/supplementary/},
\texttt{Part\_E\_Synthesis/outputs/}, and
\texttt{Part\_E\_Synthesis/figures/}.

\section{Reference Layer}

\bibliographystyle{plain}
\bibliography{universal_afl_references}
\end{document}
