\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{geometry}
\geometry{margin=2.5cm}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{criterion}{Criterion}




\title{\textbf{Paper A-12: Earth Synthesis:\\ A Constraint-Driven Flux Dynamics (CDFD) Perspective}}
\author{Steve Bico Mujjabi, MD\\
\small Independent Researcher\\
\small ORCID: 0009-0001-0556-5516}
\date{May 2026}


\newcommand{\EarthSuppPath}{Part\_A\_Earth\_Systems/supplementary/supplementary\_earth\_}
\newcommand{\EngSuppPath}{Part\_B\_Engineered\_Systems/supplementary/supplementary\_eng\_}
\newcommand{\SocSuppPath}{Part\_C\_Socioeconomic\_Systems/supplementary/supplementary\_soc\_}
\newcommand{\DomainSuppPath}{Part\_D\_Domain\_Applications/supplementary/supplementary\_dom\_}
\begin{document}
\maketitle

\begin{abstract}
This paper closes the Part A Earth-systems sequence. It treats atmosphere,
ocean, hydrology, soil, ecology, biodiversity, pollution, agriculture,
sustainability, collapse, plate tectonics, and volcanology as different
transport settings for the same CDFD/AFL audit variables: flux $\Phi$,
constraint $C$, responsiveness $S$, and structural memory $M_s$. The synthesis
does not merge these disciplines into one mechanism. It gives a common way to
ask what is flowing, what limits it, how the system responds, and what memory
keeps shaping later behavior.
\end{abstract}

\section{Series Context}

Part I sets the flow--constraint language used here \cite{MujjabiPartI2026}.
Part II develops the Life Number and the origins-of-life use of the same notation
\cite{MujjabiPartII2026}. Part III applies AFL to biology and medicine
\cite{MujjabiPartIII2026}. Part IV tests whether the notation remains useful
when it is moved into Earth systems and other domains.

\paragraph{Notation.}
In this paper, $C$ names the active Earth-system constraint or capacity burden.
$\Phi$ is the driving flow, $S$ is the response or buffering capacity, and
$M_s$ is retained structural memory. $\Lambda$ appears only where the Part II
Life Number is referenced \cite{MujjabiPartII2026}.

\section{Part A Integration}

The Part A sequence is ordered as a transport ladder. Atmospheric circulation
and ocean currents move heat and moisture at planetary scale. Hydrological and
soil systems route water, sediment, nutrients, and organic matter. Ecological,
biodiversity, pollution, and agricultural papers move into coupled biological
and human-modified transport. Sustainability and collapse ask when response
capacity is lost. Plate tectonics and volcanology extend the same audit
language to slow geophysical flow and abrupt release.

\section{Shared Earth-System Map}

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
CDFD term & Earth-system synthesis reading & Example proxy \\
\midrule
$\Phi$ & environmental drive or throughput & heat, runoff, sediment, biomass, magma \\
$C$ & active constraint & aridity, stratification, toxicity, friction, rigidity \\
$S$ & response capacity & mixing, buffering, biodiversity, storage, rerouting \\
$M_s$ & retained memory & hysteresis, degradation, legacy load, thermal storage \\
$\Psi_s$ & operating ratio & balanced, constrained, overloaded, or locked regime \\
\bottomrule
\end{tabular}
\end{table}

\section{What Carries Across Domains}

The common feature is not material identity. Heat transport, river discharge,
ecosystem energy flow, pollution load, tectonic strain, and volcanic pressure
are physically different. What carries across Part A is the audit question:
does rising flux meet a constraint that cannot relax quickly enough, and does
prior load make recovery harder? When that question can be written with real
proxies, CDFD/AFL gives a compact measurement frame.

\section{Measurement Layer}

Each Earth-system paper needs a named dataset or field measurement before the
mapping becomes more than a modelling claim. The synthesis layer asks for four
items: an observable $\Phi$, an observable $C$, a response proxy $S$, and a
memory proxy $M_s$. The failure condition can be flood, stagnation, biodiversity
loss, contamination persistence, transport weakening, or another domain-specific
state. The release does not turn those candidates into operational forecasts.

\section{Current Runtime Diagnostic}

The Part E diagnostic run includes Earth-system adapter rows for climate and
atmosphere, ocean transport, hydrology, biodiversity, and pollution. The
universal cascade stress test checks finite behavior and memory locking in the
current CDFD Runtime \cite{CDFDRuntime2026}. The numbers are a reproducibility
record for the release, not field validation.

\section{Tests That Can Break the Mapping}

The Part A synthesis weakens if the same datasets are explained more cleanly by
standard domain models, if CDFD/AFL variables cannot be separated from ordinary
descriptive labels, or if the proposed operating ratio fails to track overload
or recovery across independent Earth-system cases.

\section{Conclusion}

Part A is strongest when it stays close to measurement. Its contribution is a
shared language for flow, constraint, response, and memory across Earth systems.
The next step is not broader analogy; it is narrower data work in each domain.

\section*{Conflict of Interest}
The author declares no conflict of interest.

\section*{Data Availability}
Release-local runtime diagnostics are available under each Part's
\texttt{outputs/} folder, with release-wide synthesis artifacts under
\texttt{Part\_E\_Synthesis/supplementary/},
\texttt{Part\_E\_Synthesis/outputs/}, and
\texttt{Part\_E\_Synthesis/figures/}.

\section{Reference Layer}

\bibliographystyle{plain}
\bibliography{universal_afl_references}
\end{document}
