\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{setspace}
\onehalfspacing

\title{\textbf{Paper A-13: Earth Plate Tectonics:\\ A Constraint-Driven Flux Dynamics (CDFD) Perspective}}
\author{Steve Bico Mujjabi, MD\\
\small Independent Researcher\\
\small ORCID: 0009-0001-0556-5516}
\date{May 2026}

\begin{document}
\maketitle


\begin{abstract}
This paper asks whether Constraint-Driven Flux Dynamics (CDFD) and Adaptive Flux Limitation (AFL) can give the named system a sharper audit language. The working variables are driving flux $\Phi$, constraint $C$, surface responsiveness $S$, and structural memory $M_s$. The useful question is plain: what can be measured, what would count as overload, and what result would make the mapping fail?
\end{abstract}

\section{Series Context}

Part I sets the flow--constraint language used here \cite{MujjabiPartI2026}. Part II develops the Life Number and the origins-of-life use of the same notation \cite{MujjabiPartII2026}. Part III applies AFL to biology and medicine \cite{MujjabiPartIII2026}. Part IV is the cross-domain stress test: it asks where the notation clarifies measurement, and where it becomes only a metaphor.

\section{Introduction}

\paragraph{Notation.} In Part IV, $C$ names the active constraint or capacity burden in the system being discussed. $\Phi$ is the driving flow, $S$ is the surface response, and $M_s$ is retained structural memory. $\Lambda$ appears only where the Part II Life Number is being referenced \cite{MujjabiPartII2026}. Other Greek symbols are local coefficients whose meaning is set by the equation in which they appear.

\subsection{The Mujjabi Universal Capacity Law}
Building directly upon the Fundamental Physics (Part I) \cite{MujjabiPartI2026}, the \textbf{Mujjabi Universal Capacity Law} proposes that across all scales---from black hole event horizons to global financial markets and power grids---driving high flux ($\Phi$) against a rigid topological constraint ($C$) can drive system saturation. When $\Phi/C$ approaches critical thresholds, the system does not scale linearly; it undergoes a topological phase transition.

\subsection{The Mujjabi Universal Memory Law}
The \textbf{Mujjabi Universal Memory Law} frames persistent systemic collapse. When a complex network fails to clear its structural memory ($M_s$) and its relaxation parameter decays ($\tau_{relax} \to \infty$), temporary stressors become permanent rigidities. This is the proposed CDFD mechanism behind ecological death, economic depressions, and infrastructural gridlock.

The Earth's crust is not static; it is a continuously evolving mosaic of rigid plates floating atop a viscous, convecting mantle. Where these plates meet, immense stress accumulates along fault zones. Earthquakes occur when the accumulated stress exceeds the frictional strength of the fault, resulting in a sudden, violent displacement.

In CDFD, tectonic fault lines are not merely mechanical interfaces; they are dynamic regulatory constraints across a massive thermodynamic network. Like a clogged biological artery or a bottlenecked supply chain, a locked fault is a system struggling to process its requisite throughput.

\section{The Tectonic Adaptive Surface}
We govern the fault mechanics using the Universal Adaptive Ratio:
\begin{equation}
    \Psi_s = \left( \frac{\Phi}{C} \right) \cdot S \cdot M_s
\end{equation}

In the context of plate boundaries:
\begin{itemize}
    \item \textbf{Flux ($\Phi$)}: The continuous kinetic driving force generated by mantle convection and slab pull pushing the tectonic plates together or past one another.
    \item \textbf{Constraint ($C$)}: The frictional resistance of the fault line, specifically the interlocking rock asperities that prevent smooth sliding.
    \item \textbf{Responsiveness ($S$)}: The elastic capacity of the surrounding crust to deform and absorb stress without breaking.
    \item \textbf{Memory ($M_s$)}: Fault locking. As the fault remains stationary over centuries, chemical bonding and mechanical interlocking fuse the rock faces together. The fault "remembers" its static position, driving $M_s \to 1.0$ and cementing the constraint.
\end{itemize}

\subsection{The Seismic Cycle as a Regime Shift}
During the interseismic period, the tectonic flux ($\Phi$) continues unceasingly from the mantle. However, because the fault is locked by its memory ($M_s$) and friction ($C$), the physical flow of the crust halts at the boundary. The system enters a state of extreme localized constraint ($\Psi_s \ll 1$).

To avoid total stagnation, the system's responsiveness ($S$) flexes: the crust elastically bends, storing the incoming flux $\Phi$ as potential energy. Over decades, this elastic capacity $S$ approaches its ultimate physical limit (the yield strength of the rock).

When $S$ saturates, the adaptive ratio crashes. The system can no longer sustain the incoming flux. The stored topological tension causes a sudden, catastrophic failure of the memory term $M_s$. The interlocking asperities snap. The constraint $C$ instantly drops to near zero (dynamic sliding friction), and the stored potential energy violently converts back into kinetic flux $\Phi$---an earthquake.



\section{Measurement Layer}

The first job in any domain is to choose proxies before drawing conclusions. $\Phi$ may be a rate of material, energy, information, capital, or signal flow. $C$ is the bottleneck that slows or redirects that flow. $S$ measures the system's ability to reconfigure under stress, and $M_s$ records the past load that still changes present behavior. A useful application gives those four terms observable definitions, a time scale, and a failure condition. If a standard domain model explains the same data more cleanly, the CDFD/AFL mapping should give way.

\subsection{Current Runtime Diagnostic: Universal Network Cascade}
The diagnostic script \texttt{Part\_E\_Synthesis/supplementary/run\_partiv\_discovery.py} keeps this paper tied to the current CDFD Runtime \cite{CDFDRuntime2026}, including RK4 field updates, surface dynamics, structural-memory diffusion, and a finite-value audit. In the May 2026 run, the localized hub-drive stress test ended with hub $\Psi_s=14.8$, peak network $\Psi_s=14.8$, 21 nodes above $\Psi_s>1.5$, and 37 maximum memory-locked nodes. I treat those numbers as a runtime sanity check and as a target for later domain measurement, not as a substitute for field data.


\section{Conclusion}
Earthquakes are mathematically indistinguishable from stellar core collapse or metabolic phase shifts. They represent the abrupt topological resetting of an over-constrained system forced to shed its chronic memory to restore a sustainable flow regime. By applying CDFD, seismology transitions from a descriptive science of elastic rebound into a universal thermodynamic physics.
\section*{Conflict of Interest} The author declares no conflict of interest.
\section*{Data Availability} Release-local runtime diagnostics are available under each Part's \texttt{outputs/} folder, with release-wide synthesis artifacts under \texttt{Part\_E\_Synthesis/supplementary/}, \texttt{Part\_E\_Synthesis/outputs/}, and \texttt{Part\_E\_Synthesis/figures/}.

\section{Reference Layer}
\bibliographystyle{plain}
\bibliography{universal_afl_references}
\end{document}
