\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{geometry}
\geometry{margin=2.5cm}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{criterion}{Criterion}




\title{\textbf{Paper B-05: Neural Computation:\\ A Constraint-Driven Flux Dynamics (CDFD) Perspective}}
\author{Steve Bico Mujjabi, MD\\
\small Independent Researcher\\
\small ORCID: 0009-0001-0556-5516}
\date{May 2026}


\newcommand{\EarthSuppPath}{Part\_A\_Earth\_Systems/supplementary/supplementary\_earth\_}
\newcommand{\EngSuppPath}{Part\_B\_Engineered\_Systems/supplementary/supplementary\_eng\_}
\newcommand{\SocSuppPath}{Part\_C\_Socioeconomic\_Systems/supplementary/supplementary\_soc\_}
\newcommand{\DomainSuppPath}{Part\_D\_Domain\_Applications/supplementary/supplementary\_dom\_}
\begin{document}
\maketitle

\begin{abstract}
This paper treats neural computation as an engineered and biological transport
analogy for CDFD/AFL. Spikes, synaptic current, activation flow, gradients, and
attention-like routing are read as candidate fluxes $\Phi$ moving through
constraints $C$ such as bandwidth, inhibition, metabolic cost, finite precision,
and connectivity. The paper does not claim that biological brains and artificial
networks are the same material system. It asks which measurable bottlenecks and
memory terms make neural computation auditable.
\end{abstract}

\section{Series Context}

Part I sets the flow--constraint language used here \cite{MujjabiPartI2026}.
Part II develops the Life Number and the origins-of-life use of the same notation
\cite{MujjabiPartII2026}. Part III applies AFL to biology and medicine
\cite{MujjabiPartIII2026}. This paper sits in Part B because it focuses on the
engineered-computation reading of neural systems.

\paragraph{Notation.}
In this paper, $C$ names the active computational constraint or capacity burden.
$\Phi$ is the signal or update flow, $S$ is adaptive response, and $M_s$ is
retained structural memory. $\Lambda$ is used here only when referencing the
Part II Life Number \cite{MujjabiPartII2026}; it is not used as a claim about
consciousness or machine life.

\section{Neural Computation as Transport}

Neural computation moves signals through constrained pathways. In biological
tissue, those constraints include synaptic strength, inhibition, refractory
periods, metabolic cost, and network topology. In artificial neural networks,
they include parameter count, activation bandwidth, context length, memory
traffic, precision, optimization state, and hardware limits. CDFD/AFL gives a
single audit question for both settings: where does flow concentrate, what
limits it, how does the system adapt, and what past load remains in the
structure?

\section{Candidate Variable Map}

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
CDFD term & Neural-computation reading & Example proxy \\
\midrule
$\Phi$ & signal, activation, or gradient flow & spike rate, activation norm, token updates \\
$C$ & active constraint & inhibition, bandwidth, precision, memory traffic \\
$S$ & responsiveness & plasticity, gating, normalization, routing adjustment \\
$M_s$ & structural memory & synaptic weights, trained parameters, cached state \\
$\Psi_s$ & operating ratio & balanced, saturated, unstable, or underused regime \\
\bottomrule
\end{tabular}
\end{table}

\section{Biological and Engineered Boundaries}

Part III gives the biological and medical context for AFL \cite{MujjabiPartIII2026}.
This paper uses that context carefully. A biological neuron, a neuromorphic
device, and a transformer layer can all be described as constrained signal
systems, but the measurements differ. A useful comparison must state the unit
of flow, the time scale, the constraint, and the failure mode before drawing
any cross-domain conclusion.

\section{Measurement Layer}

For biological data, measurable proxies may include firing rate, synaptic
current, local field power, metabolic demand, refractory recovery, and plastic
change. For artificial systems, proxies may include activation magnitude,
gradient norm, attention entropy, memory bandwidth, context pressure, and
inference latency. CDFD/AFL adds value only if these proxies improve the audit
of saturation, collapse, recovery, or adaptation.

\section{Current Runtime Diagnostic}

The Part B adapter sweep includes artificial-intelligence and computation rows,
and the Part E cascade test records finite runtime behavior in the current CDFD
Runtime \cite{CDFDRuntime2026}. These files document model behavior. They do
not validate a neuroscience claim or certify an AI-system claim.

\section{Tests That Can Break the Mapping}

The mapping weakens if measured bottlenecks do not track saturation or recovery,
if ordinary neural-network or neuroscience metrics explain the behavior more
cleanly, or if the proposed variables cannot be measured without circular
definitions.

\section{Conclusion}

Neural computation is a useful Part B case because it sits between biological
signal processing and engineered computation. The CDFD/AFL frame is strongest
when it identifies real bottlenecks and memory terms rather than replacing
existing domain models.

\section*{Conflict of Interest}
The author declares no conflict of interest.

\section*{Data Availability}
Release-local runtime diagnostics are available under each Part's
\texttt{outputs/} folder, with release-wide synthesis artifacts under
\texttt{Part\_E\_Synthesis/supplementary/},
\texttt{Part\_E\_Synthesis/outputs/}, and
\texttt{Part\_E\_Synthesis/figures/}.

\section{Reference Layer}

\bibliographystyle{plain}
\bibliography{universal_afl_references}
\end{document}
