\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{geometry}
\geometry{margin=2.5cm}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}

\title{\textbf{Paper E-01: Universal AFL Synthesis:\\ A Constraint-Driven Flux Dynamics (CDFD) Perspective}}
\author{Steve Bico Mujjabi, MD\\
\small Independent Researcher\\
\small ORCID: 0009-0001-0556-5516}
\date{May 2026}

\begin{document}
\maketitle

\begin{abstract}
This synthesis states the common Part IV modelling relation and the release
boundary for cross-domain claims. Across Earth systems, engineered systems,
socioeconomic systems, domain applications, cosmic and subatomic systems, and
abstract or cognitive systems, CDFD/AFL uses the same candidate variables:
driving flux $\Phi$, constraint $C$, responsiveness $S$, and structural memory
$M_s$. Shared notation does not imply shared material mechanism or empirical
validation. The release-local runtime outputs are internal diagnostics and
tests that could break the mapping.
\end{abstract}

\paragraph{Greek-symbol convention.}
Unless a manuscript states a tighter equation-local meaning, Greek symbols are
local CDFD variables or coefficients, not universal constants. Here $\Phi$
denotes driving flux, $\Psi_s$ denotes the adaptive operating ratio, and
$\Lambda$ denotes the Life Number where used. The coefficients $\alpha$,
$\beta$, and $\gamma$ denote accumulation or reinforcement, relaxation or decay,
and diffusion, coupling, or redistribution.

\section{Common Relation}

Part IV uses the operating ratio
\begin{equation}
    \Psi_s^* = \left(\frac{\Phi}{C}\right) \cdot S \cdot M_s,
\end{equation}
with the reduced form $\Phi/C$ used only when $S=1$ and $M_s=1$. The companion
constraint-update form is
\begin{equation}
    \frac{\partial C}{\partial t} = \alpha |\Phi| - \beta C + \gamma \nabla^2 C.
\end{equation}

These equations are a modelling grammar. A domain claim becomes stronger only
when each term has a measurable proxy, a time scale, a spatial or network
scale, and a failure condition.

\section{Cross-Domain Reading}

The Part IV sections use the same form for different reasons:
\begin{itemize}
    \item Part A asks how environmental systems carry heat, water, matter, and
    biological throughput under constraint.
    \item Part B asks how engineered systems balance throughput against
    capacity, state, and responsiveness.
    \item Part C asks how social and economic systems accumulate memory,
    friction, and institutional constraint.
    \item Part D tests additional domain mappings, from immunology and
    population genetics to information flow.
    \item Part F applies the notation to cosmic and subatomic examples as
    hypothesis-level analogies.
    \item Part G treats language, science, and AI as cognitive or abstract
    transport systems.
\end{itemize}

\section{Measurement Layer}

The first job in any domain is to choose proxies before drawing conclusions. $\Phi$ may be a rate of material, energy, information, capital, or signal flow. $C$ is the bottleneck that slows or redirects that flow. $S$ measures the system's ability to reconfigure under stress, and $M_s$ records the past load that still changes present behavior. A useful application gives those four terms observable definitions, a time scale, and a failure condition. If a standard domain model explains the same data more cleanly, the CDFD/AFL mapping should give way.

\section{Current Runtime Diagnostic}

The diagnostic script \texttt{Part\_E\_Synthesis/supplementary/run\_partiv\_discovery.py}
produces a domain adapter sweep, a universal cascade stress test, two figures,
and a generated Markdown summary. The run records finite-value behavior and
overload/memory-locking dynamics in the current CDFD Runtime. It is a model
record for later measurement, not a claim that the domains have already been
validated.

\section{Claim Standard}

A Part IV claim is stronger when it names the measured $\Phi$, $C$, $S$, and
$M_s$ proxies; states the expected direction of change; gives the time and
network scale; identifies a dataset or experiment; and states what result would
falsify the CDFD/AFL interpretation.

\section{Boundary}

The universal synthesis is not operational advice for infrastructure, finance,
medicine, climate intervention, AI safety, or policy. It is a modelling archive
with reproducible local diagnostics. Any applied use requires ordinary domain
validation and expert review.

\section{Conclusion}

Part IV presents a single cross-domain modelling language. Its value depends on
whether that language produces clearer measurable hypotheses than analogy alone.
The release therefore keeps claim status, reproducibility, and falsification
standards visible beside the manuscripts and outputs.

\section*{Conflict of Interest}
The author declares no conflict of interest.

\section*{Data Availability}
Release-local runtime diagnostics are available under each Part's \texttt{outputs/} folder, with release-wide synthesis artifacts under \texttt{Part\_E\_Synthesis/supplementary/}, \texttt{Part\_E\_Synthesis/outputs/}, and \texttt{Part\_E\_Synthesis/figures/}.

\section{Reference Layer}
\bibliographystyle{plain}
\bibliography{universal_afl_references}
\end{document}
