\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{setspace}
\onehalfspacing

\title{\textbf{Paper D-01: Neural Consciousness:\\ A Constraint-Driven Flux Dynamics (CDFD) Perspective}}
\author{Steve Bico Mujjabi, MD\\
\small Independent Researcher\\
\small ORCID: 0009-0001-0556-5516}
\date{May 2026}

\begin{document}
\maketitle


\begin{abstract}
This paper asks whether Constraint-Driven Flux Dynamics (CDFD) and Adaptive Flux Limitation (AFL) can give the named system a sharper audit language. The working variables are driving flux $\Phi$, constraint $C$, surface responsiveness $S$, and structural memory $M_s$. The useful question is plain: what can be measured, what would count as overload, and what result would make the mapping fail?
\end{abstract}

\section{Series Context}

Part I sets the flow--constraint language used here \cite{MujjabiPartI2026}. Part II develops the Life Number and the origins-of-life use of the same notation \cite{MujjabiPartII2026}. Part III applies AFL to biology and medicine \cite{MujjabiPartIII2026}. Part IV is the cross-domain stress test: it asks where the notation clarifies measurement, and where it becomes only a metaphor.

\section{Introduction}

\paragraph{Notation.} In Part IV, $C$ names the active constraint or capacity burden in the system being discussed. $\Phi$ is the driving flow, $S$ is the surface response, and $M_s$ is retained structural memory. $\Lambda$ appears only where the Part II Life Number is being referenced \cite{MujjabiPartII2026}. Other Greek symbols are local coefficients whose meaning is set by the equation in which they appear.

\subsection{The Mujjabi Universal Capacity Law}
Building directly upon the Fundamental Physics (Part I) \cite{MujjabiPartI2026}, the \textbf{Mujjabi Universal Capacity Law} proposes that across all scales---from black hole event horizons to global financial markets and power grids---driving high flux ($\Phi$) against a rigid topological constraint ($C$) can drive system saturation. When $\Phi/C$ approaches critical thresholds, the system does not scale linearly; it undergoes a topological phase transition.

\subsection{The Mujjabi Universal Memory Law}
The \textbf{Mujjabi Universal Memory Law} frames persistent systemic collapse. When a complex network fails to clear its structural memory ($M_s$) and its relaxation parameter decays ($\tau_{relax} \to \infty$), temporary stressors become permanent rigidities. This is the proposed CDFD mechanism behind ecological death, economic depressions, and infrastructural gridlock.

A human brain consumes 20\% of the body's metabolic energy despite representing only 2\% of its mass. This massive energy expenditure is required to maintain the continuous electrical and chemical flux across billions of synaptic junctions. The brain is not a static computer; it is a highly turbulent fluid of information.

To prevent this immense flux from spiraling into chaotic seizures, the brain deploys massive inhibitory constraints. The interplay between excitatory transmission and inhibitory gating creates the highly ordered, complex patterns we experience as cognition.

\section{The Cognitive Adaptive Surface}
We apply the Universal Adaptive Ratio to neurobiology:
\begin{equation}
    \Psi_s = \left( \frac{\Phi}{C} \right) \cdot S \cdot M_s
\end{equation}

For a specific cortical region:
\begin{itemize}
    \item \textbf{Flux ($\Phi$)}: The rate of excitatory action potentials (e.g., glutamate transmission) and the volume of incoming sensory data.
    \item \textbf{Constraint ($C$)}: The inhibitory tone of the network (e.g., GABAergic interneurons) and the metabolic limits of the astrocytes providing energy.
    \item \textbf{Responsiveness ($S$)}: Neuroplasticity---the capacity of the tissue to rapidly grow new dendritic spines or upregulate receptor density to absorb new information.
    \item \textbf{Memory ($M_s$)}: Long-Term Potentiation (LTP) and structural engrams. The physical reinforcement of a synapse "remembers" past flux, acting as a chronic, semi-permanent modifier to the local topological constraints.
\end{itemize}

\subsection{Epilepsy and PTSD as Topological Failures}
A healthy, conscious brain operates exactly at the edge of criticality ($\Psi_s \approx 1$). The flux is maximal, but perfectly contained by the constraints, allowing complex, ordered thought to propagate.

If the inhibitory constraint $C$ drops abruptly (e.g., due to a chemical imbalance or trauma), the excitatory flux $\Phi$ runs completely unchecked. The adaptive ratio skyrockets ($\Psi_s \gg 1$). The localized region of the brain is overwhelmed, and the unchecked electrical cascade spreads topologically across the cortex. This is an \textbf{Epileptic Seizure}, the neurological equivalent of a supernova or a power grid cascading failure.

Conversely, consider Psychological Trauma (PTSD). A massively terrifying sensory input causes the brain to aggressively enforce a memory lock ($M_s \to 1.0$) on specific amygdala-hippocampal circuits. The brain permanently elevates the constraint $C$ around that specific informational pathway to prevent the traumatic flux from recurring. The local topology becomes rigidly frozen ($\Psi_s \ll 1$). The patient loses cognitive flexibility because the structural memory refuses to adapt to new, safe environments.



\section{Measurement Layer}

The first job in any domain is to choose proxies before drawing conclusions. $\Phi$ may be a rate of material, energy, information, capital, or signal flow. $C$ is the bottleneck that slows or redirects that flow. $S$ measures the system's ability to reconfigure under stress, and $M_s$ records the past load that still changes present behavior. A useful application gives those four terms observable definitions, a time scale, and a failure condition. If a standard domain model explains the same data more cleanly, the CDFD/AFL mapping should give way.

\subsection{Current Runtime Diagnostic: Universal Network Cascade}
The diagnostic script \texttt{Part\_E\_Synthesis/supplementary/run\_partiv\_discovery.py} keeps this paper tied to the current CDFD Runtime \cite{CDFDRuntime2026}, including RK4 field updates, surface dynamics, structural-memory diffusion, and a finite-value audit. In the May 2026 run, the localized hub-drive stress test ended with hub $\Psi_s=14.8$, peak network $\Psi_s=14.8$, 21 nodes above $\Psi_s>1.5$, and 37 maximum memory-locked nodes. I treat those numbers as a runtime sanity check and as a target for later domain measurement, not as a substitute for field data.


\section{Conclusion}
Consciousness is the experiential artifact of a system actively regulating its own information flow at the edge of thermodynamic criticality. By understanding the brain as an active CDFD surface, psychiatric and neurological disorders cease to be arbitrary chemical imbalances and instead reveal themselves as highly predictable topological phase failures.
\section*{Conflict of Interest} The author declares no conflict of interest.
\section*{Data Availability} Release-local runtime diagnostics are available under each Part's \texttt{outputs/} folder, with release-wide synthesis artifacts under \texttt{Part\_E\_Synthesis/supplementary/}, \texttt{Part\_E\_Synthesis/outputs/}, and \texttt{Part\_E\_Synthesis/figures/}.

\section{Reference Layer}
\bibliographystyle{plain}
\bibliography{universal_afl_references}
\end{document}
