\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{geometry}
\geometry{margin=2.5cm}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{criterion}{Criterion}




\title{\textbf{Paper D-09: AI Neural Computation:\\ A Constraint-Driven Flux Dynamics (CDFD) Perspective}}
\author{Steve Bico Mujjabi, MD\\
\small Independent Researcher\\
\small ORCID: 0009-0001-0556-5516}
\date{May 2026}


\newcommand{\EarthSuppPath}{Part\_A\_Earth\_Systems/supplementary/supplementary\_earth\_}
\newcommand{\EngSuppPath}{Part\_B\_Engineered\_Systems/supplementary/supplementary\_eng\_}
\newcommand{\SocSuppPath}{Part\_C\_Socioeconomic\_Systems/supplementary/supplementary\_soc\_}
\newcommand{\DomainSuppPath}{Part\_D\_Domain\_Applications/supplementary/supplementary\_dom\_}
\begin{document}
\maketitle

\begin{abstract}
This paper treats AI neural computation as a CDFD/AFL domain-application
hypothesis. Token flow, activation flow, gradient updates, memory retrieval,
tool calls, and inference demand are read as candidate fluxes $\Phi$ moving
through constraints $C$ such as context length, compute budget, memory bandwidth,
latency, alignment filters, and training-data structure. The goal is to define
measurable overload and recovery variables, not to claim that an AI system has
life, intention, or consciousness.
\end{abstract}

\section{Series Context}

Part I sets the flow--constraint language used here \cite{MujjabiPartI2026}.
Part II develops the Life Number and the origins-of-life use of the same notation
\cite{MujjabiPartII2026}. Part III applies AFL to biology and medicine
\cite{MujjabiPartIII2026}. This paper uses those references for notation and
series continuity; it does not transfer biological claims onto AI systems.

\paragraph{Notation.}
In this paper, $C$ names the active AI-system constraint or capacity burden.
$\Phi$ is the information or compute flow, $S$ is adaptive response, and $M_s$
is retained structural memory. $\Lambda$ is named only as the Part II Life
Number \cite{MujjabiPartII2026}. It is not used here as a machine-life variable.

\section{AI Computation as Constrained Flow}

Modern AI systems move information through layered constraints. A prompt pushes
tokens into a context window; attention routes signals through a trained
parameter field; retrieval systems add external memory; inference hardware sets
latency, bandwidth, and energy limits; safety filters and tool policies add
additional routing constraints. CDFD/AFL reads the system as a flow problem with
memory: outputs depend on current demand and on structure laid down during
training, fine-tuning, caching, and retrieval.

\section{Candidate Variable Map}

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
CDFD term & AI-system reading & Example proxy \\
\midrule
$\Phi$ & inference or update flow & token rate, request load, gradient norm \\
$C$ & active constraint & context length, memory bandwidth, latency, policy filter \\
$S$ & responsiveness & routing change, retrieval update, tool selection, scaling \\
$M_s$ & structural memory & weights, embeddings, cache, retrieval corpus, history \\
$\Psi_s$ & operating ratio & stable, saturated, brittle, or overloaded regime \\
\bottomrule
\end{tabular}
\end{table}

\section{Failure Modes}

AI overload does not need mystical language. It can mean context saturation,
retrieval conflict, latency collapse, hallucination under sparse grounding,
memory contamination, brittle tool routing, or training-inference mismatch.
Those states are measurable only when the system boundary is clear. A local
model, a hosted model, an agent loop, and a retrieval-augmented system have
different $\Phi$, $C$, $S$, and $M_s$ proxies.

\section{Measurement Layer}

Useful measurements include token throughput, context occupancy, retrieval hit
rate, cache age, latency, refusal/error rate, grounding score, tool-call
success, memory bandwidth, gradient stability, and energy cost. A CDFD/AFL
analysis must name the failure condition before interpreting $\Psi_s$. Without
that condition, the notation only renames ordinary engineering metrics.

\section{Current Runtime Diagnostic}

The Part B adapter sweep includes an artificial-intelligence row, and the Part
E cascade diagnostic checks finite behavior and memory locking in the current
CDFD Runtime \cite{CDFDRuntime2026}. These outputs are release evidence for the
runtime, not evidence that any deployed AI system has been validated by this
paper.

\section{Tests That Can Break the Mapping}

The AI mapping weakens if the proposed variables cannot predict saturation,
latency, grounding failure, or recovery better than existing system metrics; if
$M_s$ cannot be separated from ordinary model state; or if the analysis depends
on anthropomorphic language instead of measurable system behavior.

\section{Conclusion}

AI neural computation is a strong stress test for Part IV because it has visible
flow, constraint, responsiveness, and memory. The claim stays modest: CDFD/AFL
may help audit overload and recovery in AI systems when the variables are
measured plainly.

\section*{Conflict of Interest}
The author declares no conflict of interest.

\section*{Data Availability}
Release-local runtime diagnostics are available under each Part's
\texttt{outputs/} folder, with release-wide synthesis artifacts under
\texttt{Part\_E\_Synthesis/supplementary/},
\texttt{Part\_E\_Synthesis/outputs/}, and
\texttt{Part\_E\_Synthesis/figures/}.

\section{Reference Layer}

\bibliographystyle{plain}
\bibliography{universal_afl_references}
\end{document}
