\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{geometry}
\geometry{margin=2.5cm}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}
\newtheorem{theorem}{Theorem}

\title{\textbf{Paper D-10: Information Flow:\\ A Constraint-Driven Flux Dynamics (CDFD) Perspective}}
\author{Steve Bico Mujjabi, MD\\
\small Independent Researcher\\
\small ORCID: 0009-0001-0556-5516}
\date{May 2026}

\begin{document}
\maketitle

\begin{abstract}
This paper treats digital information flow as a CDFD/AFL modelling hypothesis.
Posts, shares, links, semantic updates, and attention are represented as
candidate fluxes $\Phi$ moving through cognitive, algorithmic, and network
constraints $C$. The goal is not to certify a social-media model. The goal is to
state measurable variables and failure conditions for overload, echo chambers,
viral cascades, and semantic bottlenecks. Release-local runtime outputs are
internal diagnostics and tests that could break the mapping, not field evidence.
\end{abstract}

\section{Series Context}

Part I sets the flow--constraint language used here \cite{MujjabiPartI2026}.
Part II develops the Life Number and the origins-of-life use of the same notation
\cite{MujjabiPartII2026}. Part III applies AFL to biology and medicine
\cite{MujjabiPartIII2026}. This paper uses the series notation for information
flow while keeping social and platform claims at hypothesis level.

\paragraph{Greek-symbol convention.}
Unless a manuscript states a tighter equation-local meaning, Greek symbols are
local CDFD variables or coefficients, not universal constants. Here $\Phi$
denotes driving flux, $\Psi_s$ denotes the adaptive operating ratio, and
$\Lambda$ denotes the Life Number where used \cite{MujjabiPartII2026}. The coefficients $\alpha$,
$\beta$, and $\gamma$ denote accumulation or reinforcement, relaxation or decay,
and diffusion, coupling, or redistribution.

\section{Information as Constrained Transport}

Digital networks move information quickly, but not freely. Human attention,
platform ranking, moderation queues, recommendation systems, language
boundaries, and trust relationships all act as constraints. CDFD/AFL frames the
system as a transport problem: when information flux rises faster than attention
and routing constraints can relax, the system can enter overload, fragmentation,
or repetitive circulation.

\section{Candidate Variable Map}

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
CDFD term & Information-flow reading & Example proxy \\
\midrule
$\Phi$ & semantic or attention flux & posts, shares, impressions, messages \\
$C$ & attention or routing constraint & ranking friction, moderation delay, trust barrier \\
$S$ & responsiveness & fact checking, topic diversity, cross-cluster routing \\
$M_s$ & memory & prior beliefs, platform history, cached ranking state \\
$\Psi_s$ & operating ratio & overload, virality, isolation, or balanced flow \\
\bottomrule
\end{tabular}
\end{table}

\section{Echo Chambers and Viral Cascades}

An echo chamber is treated here as a candidate low-constraint loop inside a
larger high-constraint network. A viral cascade is treated as a local flux surge
that exceeds available attention and routing capacity. These are modelling
hypotheses. They need platform data, time scales, and comparison against
standard network-science baselines before they become empirical claims.

\section{Measurement Layer}

The first job in any domain is to choose proxies before drawing conclusions. $\Phi$ may be a rate of material, energy, information, capital, or signal flow. $C$ is the bottleneck that slows or redirects that flow. $S$ measures the system's ability to reconfigure under stress, and $M_s$ records the past load that still changes present behavior. A useful application gives those four terms observable definitions, a time scale, and a failure condition. If a standard domain model explains the same data more cleanly, the CDFD/AFL mapping should give way.

\section{Current Runtime Diagnostic}

The release-local domain adapter sweep includes an information-theory row and
related network-routing rows. Those artifacts test whether the current runtime
returns finite outputs and interpretable regimes under toy conditions. They do
not validate polarization, virality, or platform-behavior claims.

\section{Falsification Targets}

The information-flow mapping weakens if measured attention and routing
constraints do not track overload, if echo-chamber formation is better explained
without CDFD/AFL variables, or if the model cannot distinguish between semantic
content, network topology, and platform policy effects.

\section{Conclusion}

Information flow can be studied as constrained transport when the variables are
defined carefully. In Part IV, the CDFD/AFL contribution is an audit language
for candidate overload and memory effects, not a proof of social behavior.

\section*{Conflict of Interest}
The author declares no conflict of interest.

\section*{Data Availability}
Release-local runtime diagnostics are available under each Part's \texttt{outputs/} folder, with release-wide synthesis artifacts under \texttt{Part\_E\_Synthesis/supplementary/}, \texttt{Part\_E\_Synthesis/outputs/}, and \texttt{Part\_E\_Synthesis/figures/}.

\section{Reference Layer}
\bibliographystyle{plain}
\bibliography{universal_afl_references}
\end{document}
